Assessing the Firing of Ceramic Materials: a Seriation-Based Approach

### 2023-06-08 analysis/paper/paper.qmd
Add figures

---
title: "Assessing the Firing of Ceramic Materials: a Seriation-Based Approach"
author:
  - Nicolas Frerebeau:
      correspondence: "yes"
      email: nicolas.frerebeau@u-bordeaux-montaigne.fr
      orcid: 0000-0001-5759-4944
      affiliations:
        - name: "UMR 6034 Archéosciences Bordeaux"
          address: "Maison de l'Archéologie, Université Bordeaux Montaigne"
          city: Pessac cedex
          country: France
          postal-code: 33607
          url: https://www.archeosciences-bordeaux.fr
title-block-published: "Last updated"  
date: now
date-format: long
format: 
  pdf: 
    documentclass: scrartcl
    include-in-header: 
      - text: |
          \usepackage{booktabs}
    classoption: 
      - french
      - 12pt
      - a4paper
      - oneside
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: false
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: [references.bib, packages.bib]
csl: "../templates/archeosciences.csl"
abstract: |
  The study of the firing of archaeological ceramics is thus a key factor in 
  interpreting ancient techniques and related social practices. This study 
  proposes a new method based on the correspondence analysis of powder X-ray 
  diffraction data (XRD), aiming to provide an alternative to the so-called
  Equivalent Firing Temperature framework. The properties of correspondence 
  analysis and its application for matrix seriation enable the description of 
  the firing of a ceramic assemblage. The application of correspondence 
  analysis to XRD data of Iron Age ceramics revealed the  significant potential 
  of this statistical method in the technological analysis of ancient pottery. 
  This allows to order of a large number of samples according to a firing 
  gradient and to track mineralogical transformations occurring during firing.
keywords: |
  X-ray diffraction; ceramic firing; correspondence analysis; matrix seriation
---

>>>>

---
title: "Assessing the Firing of Ceramic Materials: a Seriation-Based Approach"
author:
  - Nicolas Frerebeau:
      correspondence: "yes"
      email: nicolas.frerebeau@u-bordeaux-montaigne.fr
      orcid: 0000-0001-5759-4944
      affiliations:
        - name: "UMR 6034 Archéosciences Bordeaux"
          address: "Maison de l'Archéologie, Université Bordeaux Montaigne"
          city: Pessac cedex
          country: France
          postal-code: 33607
          url: https://www.archeosciences-bordeaux.fr
title-block-published: "Last updated"  
date: now
date-format: long
format: 
  pdf: 
    documentclass: scrartcl
    include-in-header: 
      - text: |
          \usepackage{booktabs}
    classoption: 
      - french
      - 12pt
      - a4paper
      - oneside
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
knitr:
  opts_chunk:
    echo: false
    warning: false
    message: false
    comment: "#>"
    fig.path: "../figures/"
    fig.dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: [references.bib, packages.bib]
csl: "../templates/archeosciences.csl"
abstract: |
  The study of the firing of archaeological ceramics is thus a key factor in 
  interpreting ancient techniques and related social practices. This study 
  proposes a new method based on the correspondence analysis of powder X-ray 
  diffraction data (XRD), aiming to provide an alternative to the so-called
  Equivalent Firing Temperature framework. The properties of correspondence 
  analysis and its application for matrix seriation enable the description of 
  the firing of a ceramic assemblage. The application of correspondence 
  analysis to XRD data of Iron Age ceramics revealed the  significant potential 
  of this statistical method in the technological analysis of ancient pottery. 
  This allows to order of a large number of samples according to a firing 
  gradient and to track mineralogical transformations occurring during firing.
keywords: |
  X-ray diffraction; ceramic firing; correspondence analysis; matrix seriation
---
